Analysis of Massachusetts' Standardized Testing Multi-Group Structural Equation Modelling

### 2014-04-12 paper-knitr.tex
Presentation RNW

\subsection{Theoretical structure of the test}
Our paper begins with an analysis of the psychometric equivalence of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.

@@

Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions; most work under the Bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}. However, within the structure of the exam, missing data was unsurprisingly a rare occurrence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions. Even so, for the MG-CFA we attempted to account for this by utilizing the Full Information Maximum Likelihood (FIML) estimation in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors. However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques as it produces potentially biased results under the assumption that items are missing completely at random. As such, we processed the original responses using the mice package\cite{mice} using 13 imputed data frames with 15 iterations in each dataframe. Regression trees are relatively robust to missing data because all information is not processed simultaneously, but rather sequentially in order of variable importance. As such we did not take any additional steps when conducting that analysis.

\section{Results}
\subsection{Exploratory factor analyses}
Given the preceding information, we examined the Mathematics, English, and Biology sections of the exam using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains. Factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items. The psych package was used to calculate the factor structure. Items that loaded higher than .3 on a factor were specified as indicators, and diagrams of each factor can be located in Appendix A.

>>>>

Our paper begins with an analysis of the psychometric equivalence of the MCAS test within our sample of 10\textsuperscript{th} grade students, followed by an investigation into performance within each knowledge domain, conditioned upon the demographics and item responses found within a questionnaire included with the test.

@@

Missing data in survey and test data is an important consideration, and there exist a number of techniques to produce more robust solutions; most work under the Bayesian estimation process for missing data pioneered by Donald Rubin\cite{rubin}. However, within the structure of the exam, missing data was unsurprisingly a rare occurrence: the highest proportion of missing for any item was .008, and 91.23\% had no missing questions. Even so, for the MG-CFA we attempted to account for this by utilizing the Full Information Maximum Likelihood (FIML) estimation in conjunction with a large number of bootstraps to obtain accurate estimates of asymptotic errors. However, the questionnaire exhibited extreme proportions of non-response, which is problematic for the multivariate regression techniques as it produces potentially biased results under the assumption that items are missing completely at random. As such, we processed the original responses using the mice package\cite{mie} using 13 imputed data frames with 15 iterations in each dataframe. Regression trees are relatively robust to missing data because all information is not processed simultaneously, but rather sequentially in order of variable importance. As such we did not take any additional steps when conducting that analysis.

\section{Results}
\subsection{Exploratory factor analyses}
Given the preceding information, we examined the Mathematics, English, and Biology sections of the exam using a oblimin rotated factor structure, with the number of latent factors equal to the number of general domains. Factor loadings were calculated using maximum likelihood estimates from the tetrachoric correlation matrices for each subset of items. The psych package was used to calculate the factor structure. Items that loaded higher than .3 on a factor were specified as indicators. See Appendix A for diagrams of the factor structures.

### 2014-04-12 presentation.Rnw
Presentation RNW

\documentclass{beamer}

\usepackage{apacite}
\setbeamertemplate{bibliography item}[text]

<<setup, include=F, cache=FALSE, echo=FALSE>>=
load('~/upstat-competition/.RData')
require(psych)
require(semTools)
require(mvpart)
setwd("~/upstat-competition/")
Sys.setenv(TEXINPUTS=getwd(), 
           BIBINPUTS=getwd(), 
           BSTINPUTS=getwd())
@

@@

@@

  \begin{frame}[t,allowframebreaks]
    \frametitle{References}
    \nocite{*}
    \bibliography{references}
  \end{frame}

\end{document}

>>>>

\documentclass{beamer}
\usepackage{apacite,geometry,fancyhdr}

@@

<<setup, include=FALSE, cache=FALSE, echo=FALSE, results='hide'>>=
library(lavaan)
library(psych)
library(mvpart)
library(semTools)

load("comp.RData")

Sys.setenv(TEXINPUTS=getwd(), 
           BIBINPUTS=getwd(), 
           BSTINPUTS=getwd())
@

@@

\begin{frame}[allowframebreaks]
\frametitle{References}
\nocite{*}
\bibliography{references}{}
\bibliographystyle{apacite}
\end{frame}
\end{document}

### 2014-04-12 test.Rnw
Presentation RNW

\documentclass{beamer}
\usepackage{apacite,geometry,fancyhdr}

@@

    \begin{frame}{results}
      <<fig=T,echo=F>>=
      fa.diagram(feng,main='English Oblimin')
      #fa.diagram(fmat,main='Mathematics Oblimin')
      #fa.diagram(fsci1,main='Biology Oblimin')
      @
    \end{frame}

@@

\section{Conclusions}
  \begin{frame}{Conclusions}

\begin{frame}[allowframebreaks]
\frametitle{References}
\nocite{*}
\bibliography{references}{}
\bibliographystyle{apacite}
\end{frame}
\end{document}

>>>>

\documentclass{beamer}
\usepackage{apacite,geometry,fancyhdr,graphicx}

@@

    \begin{frame}{Results}
      \begin{center}
        \includegraphics[keepaspectratio=true,width=2.7in]{english}
      \end{center}
    \end{frame}

    \begin{frame}{Results}
      \includegraphics[keepaspectratio=true,width=4.3in]{mathematics}
    \end{frame}

    \begin{frame}{Results}
      \includegraphics[keepaspectratio=true,width=4.3in]{biology}
    \end{frame}

@@

\section{Conclusions}
  \begin{frame}{Conclusions}
    \begin{itemize}
      \item Fails strict invariance \pause
      \item Regression tree: split on ethnicity
    \end{itemize}
  \end{frame}

\begin{frame}[allowframebreaks]
\frametitle{References}
\nocite{*}
\bibliography{references}{}
\bibliographystyle{apacite}
\end{frame}

\begin{frame}
  \frametitle{Appendix A}
  <<fig=T, echo=FALSE, fig.height=4, fig.width=5>>=
fa.diagram(feng,main='English Oblimin')
@
\end{frame}

\begin{frame}
  \frametitle{Appendix A}
  <<fig=T, echo=FALSE, fig.height=4, fig.width=5>>=
fa.diagram(fmat,main='Mathematics Oblimin')
@
\end{frame}

\begin{frame}
  \frametitle{Appendix A}
  <<fig=T, echo=FALSE, fig.height=4, fig.width=5>>=
fa.diagram(fsci1,main='Biology Oblimin')
@
\end{frame}

\end{document}
